---
title: "Player name lookup table"
author: "David Selby"
date: '2022-06-24'
output: html_document
---

In this document we will create a lookup table for matching player names, as downloaded from the Elo ratings web site, with the names given in the submission template.

```{r}
elo_m <- read.csv('data/elo_ratings/atp22.csv')
elo_f <- read.csv('data/elo_ratings/wta22.csv')
template <- read.csv(file('submission-template.csv', encoding = 'UTF-8'))
```

Get the unique names in the two sources.

```{r}
elo_names <- c(elo_m$Player, elo_f$Player)
template_names <- union(template$player1_name, template$player2_name)
```

Which names don't match up?

```{r}
template_names[!template_names %in% tolower(elo_names)]
```

Rather than do anything clever, we can just manually fix these.
Let's first join those that match exactly (except for case).

```{r, message = F}
library(dplyr)
lookup <- distinct(template, player = player1_name, id = player1_id, Gender) %>%
  rbind(distinct(template, player = player2_name, id = player2_id, Gender)) %>%
  distinct()

elo_tbl <- data.frame(elo_name = elo_names,
                      lower = tolower(elo_names))

lookup <- lookup %>%
  left_join(elo_tbl, by = c(player = 'lower'))
```

## Fuzzy string matching

To save a bit of manual work, we can use fuzzy string matching, which (surprisingly, to me) is built into base R.

https://statisticsglobe.com/fuzzy-matching-r

```{r}
fuzzy_match <- function(name, list, max_distance = 4) {
  sapply(name,
         \(x) agrep(x, list, max.distance = max_distance, value = TRUE)
  )
}

lookup <- lookup %>%
  mutate(fuzzy = ifelse(is.na(elo_name), fuzzy_match(player, elo_names), NA),
         single = ifelse(lengths(fuzzy) > 1, NA, sapply(fuzzy, '[', 1)),
         elo_name = coalesce(elo_name, single)) %>%
  select(-single)
```

Unfortunately we can't rely on fuzzy matching for all of them because of Xinyu Wang vs. Xi Yu Wang (who are different people).

```{r}
lookup %>%
  filter(is.na(elo_name))
```
Life's short. Fix them manually.

```{r}
lookup <- lookup %>%
  mutate(elo_name = case_when(player == 'xinyu wang' ~ 'Xin Yu Wang',
                              player == 'albert ramos-vinolas' ~ 'Albert Ramos',
                              !is.na(elo_name) ~ elo_name,
                              TRUE ~ NA_character_)) %>%
  select(-fuzzy)
```

Any remaining issues?

```{r}
filter(lookup, is.na(elo_name))
```

Save to disk.

```{r}
write.csv(lookup, file('data/name_lookup.csv', encoding = 'UTF-8'),
          row.names = FALSE)
```