---
title: "Download Elo ratings"
author: "David Selby"
date: '2022-06-21'
output:
  html_document:
    default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a function that downloads the Elo ratings table from https://tennisabstract.com/reports/atp_elo_ratings.html (or WebArchive versions thereof).

```{r}
library(rvest)
download_elo <- function(url) {
  read_html(url) %>%
    html_element('#reportable') %>%
    html_table()
}
```

Example:

```{r}
atp22 <- download_elo('https://tennisabstract.com/reports/atp_elo_ratings.html')
head(atp22)
```

We can clean this up a little bit. Get rid of the empty columns.
Also, (you'll thank me for this later!) get rid of non-breaking spaces in the player names.

```{r message = F}
library(dplyr)
clean_elo <- function(tbl) {
  tbl <- tbl[, nchar(colnames(tbl)) > 0]
  mutate(tbl, across(where(is.character),
                     ~ stringr::str_replace_all(.x, '\\s', ' ')))
}

clean_elo(atp22) %>% head()

download_elo2 <- function(url)
  download_elo(url) %>% clean_elo()
```

And get some historic data from the Web Archive.

```{r}
atp21 <- download_elo2('https://web.archive.org/web/20210624053134/https://tennisabstract.com/reports/atp_elo_ratings.html')
atp19 <- download_elo2('https://web.archive.org/web/20190627232139/https://tennisabstract.com/reports/atp_elo_ratings.html')
atp18 <- download_elo2('https://web.archive.org/web/20180629003041/https://tennisabstract.com/reports/atp_elo_ratings.html')
```

I would combine these into one big table with a column for 'year', but they changed the names of the columns slightly from one year to another.
If we are interested in 'grass Elo' then it might be worth checking what the difference is between the 'raw grass' and 'gElo' scores.

```{r}
head(atp21)
head(atp19)
head(atp18)
```

Save everything to file.

```{r}
atp22 <- clean_elo(atp22)
write.csv(atp22, 'data/elo_ratings/atp22.csv', row.names = F)
write.csv(atp21, 'data/elo_ratings/atp21.csv', row.names = F)
write.csv(atp19, 'data/elo_ratings/atp19.csv', row.names = F)
write.csv(atp18, 'data/elo_ratings/atp18.csv', row.names = F)
```

Don't forget the womens' tournament!

```{r}
wta22 <- download_elo2('https://tennisabstract.com/reports/wta_elo_ratings.html')
wta21 <- download_elo2('https://web.archive.org/web/20210624053134/https://tennisabstract.com/reports/wta_elo_ratings.html')
wta19 <- download_elo2('https://web.archive.org/web/20190627232139/https://tennisabstract.com/reports/wta_elo_ratings.html')
wta18 <- download_elo2('https://web.archive.org/web/20180629003041/https://tennisabstract.com/reports/wta_elo_ratings.html')
```

```{r}
write.csv(wta22, 'data/elo_ratings/wta22.csv', row.names = F)
write.csv(wta21, 'data/elo_ratings/wta21.csv', row.names = F)
write.csv(wta19, 'data/elo_ratings/wta19.csv', row.names = F)
write.csv(wta18, 'data/elo_ratings/wta18.csv', row.names = F)
```

## Odds

We scrape the tables of betting odds from 

- https://www.oddsportal.com/tennis/united-kingdom/atp-wimbledon/
- https://www.oddsportal.com/tennis/united-kingdom/wta-wimbledon/

Paste into Excel, then save to CSV.

```{r}
men_odds <- read.csv('data/odds/men_raw1.csv', header = F)
women_odds <- read.csv('data/odds/women_raw1.csv', header = F)
```

```{r}
men_odds <- men_odds %>%
  setNames(c('time', 'pair', 'blank', 'cancelled', 'odds1', 'odds2', 'Bs')) %>%
  filter(nchar(pair) > 0,            # skip empty / meta rows
         nchar(cancelled) == 0) %>%  # delete cancelled matches
  select(-blank, -cancelled, -Bs) %>%
  tidyr::separate(pair, c('player1', 'player2'), sep = ' - ')
```

Due to cancelled matches the layout is not guaranteed to be the same for every table.

```{r}
women_odds <- women_odds %>%
  setNames(c('time', 'pair', 'blank', 'cancelled', 'odds1', 'odds2', 'Bs')) %>%
  filter(nchar(pair) > 0,
         nchar(cancelled) == 0) %>%
  select(-blank, -cancelled, -Bs) %>%
  tidyr::separate(pair, c('player1', 'player2'), sep = ' - ')
```

```{r}
write.csv(women_odds, 'data/odds/women1.csv', row.names = F)
write.csv(men_odds, 'data/odds/men1.csv', row.nams = F)
```
