---
title: "Download Elo ratings"
author: "David Selby"
date: '2022-06-21'
output:
  html_document:
    default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a function that downloads the Elo ratings table from https://tennisabstract.com/reports/atp_elo_ratings.html (or WebArchive versions thereof).

```{r}
library(rvest)
download_elo <- function(url) {
  read_html(url) %>%
    html_element('#reportable') %>%
    html_table()
}
```

Example:

```{r}
atp22 <- download_elo('https://tennisabstract.com/reports/atp_elo_ratings.html')
head(atp22)
```

We can clean this up a little bit. Get rid of the empty columns.

```{r}
clean_elo <- function(tbl) {
  tbl[, nchar(colnames(tbl)) > 0]
}

clean_elo(atp22) %>% head()

download_elo2 <- function(url)
  download_elo(url) %>% clean_elo()
```

And get some historic data from the Web Archive.

```{r}
atp21 <- download_elo2('https://web.archive.org/web/20210624053134/https://tennisabstract.com/reports/atp_elo_ratings.html')
atp19 <- download_elo2('https://web.archive.org/web/20190627232139/https://tennisabstract.com/reports/atp_elo_ratings.html')
atp18 <- download_elo2('https://web.archive.org/web/20180629003041/https://tennisabstract.com/reports/atp_elo_ratings.html')
```

I would combine these into one big table with a column for 'year', but they changed the names of the columns slightly from one year to another.
If we are interested in 'grass Elo' then it might be worth checking what the difference is between the 'raw grass' and 'gElo' scores.

```{r}
head(atp21)
head(atp19)
head(atp18)
```

Save everything to file.

```{r}
atp22 <- clean_elo(atp22)
write.csv(atp22, 'data/elo_ratings/atp22.csv', row.names = F)
write.csv(atp21, 'data/elo_ratings/atp21.csv', row.names = F)
write.csv(atp19, 'data/elo_ratings/atp19.csv', row.names = F)
write.csv(atp18, 'data/elo_ratings/atp18.csv', row.names = F)
```

Don't forget the womens' tournament!

```{r}
wta22 <- download_elo2('https://tennisabstract.com/reports/wta_elo_ratings.html')
wta21 <- download_elo2('https://web.archive.org/web/20210624053134/https://tennisabstract.com/reports/wta_elo_ratings.html')
wta19 <- download_elo2('https://web.archive.org/web/20190627232139/https://tennisabstract.com/reports/wta_elo_ratings.html')
wta18 <- download_elo2('https://web.archive.org/web/20180629003041/https://tennisabstract.com/reports/wta_elo_ratings.html')
```

```{r}
write.csv(wta22, 'data/elo_ratings/wta22.csv', row.names = F)
write.csv(wta21, 'data/elo_ratings/wta21.csv', row.names = F)
write.csv(wta19, 'data/elo_ratings/wta19.csv', row.names = F)
write.csv(wta18, 'data/elo_ratings/wta18.csv', row.names = F)
```
