---
title: "Manual name matching"
author:
  - Ian Hamilton
  - Stefan Stein
  - David Selby
date: June 2022
output:
  html_document:
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = TRUE)
```

Open the Elo ranking and odds data, find names that cannot be matched automatically, do those for which Elo rankings exist in Excel.

```{r}
library(tidyverse)
library(igraph)
```


```{r}
transform_name <- function(name){
  splitName <- str_split(name, " ")[[1]]
  
  lastName <- tail(splitName, n=1)%>%str_to_title()
  firstName <- head(splitName, -1)%>%substr(start = 1, stop = 1)%>%toupper()%>%paste(".",sep = "")%>%paste(collapse = "")
  
  return(paste(lastName,firstName))
}
```


# Men

```{r}
df_2018 <- read_csv("data/Men 2018.csv", col_types = cols(Date = col_date(format = "%d/%m/%Y")))%>%
  filter(Tournament == "Wimbledon" & Round == "1st Round")%>%
  select(Winner, Loser, AvgW, AvgL)%>%
  mutate(p_ij = 1 / AvgW,
         p_ji = 1/ AvgL,
         total_prob = p_ij + p_ji) %>%
  mutate(p_ij = p_ij / total_prob,
         p_ji = p_ji / total_prob) %>%
  select(-total_prob) %>%
  mutate(logit_p_ij = log(p_ij / p_ji),
         lambda_w = 0.5*logit_p_ij,
         lambda_l = -0.5*logit_p_ij)

df_2019 <- read_csv("data/Men 2019.csv", col_types = cols(Date = col_date(format = "%d/%m/%Y")))%>%
  filter(Tournament == "Wimbledon" & Round == "1st Round")%>%
  select(Winner, Loser, AvgW, AvgL)%>%
  mutate(p_ij = 1 / AvgW,
         p_ji = 1/ AvgL,
         total_prob = p_ij + p_ji) %>%
  mutate(p_ij = p_ij / total_prob,
         p_ji = p_ji / total_prob) %>%
  select(-total_prob) %>%
  mutate(logit_p_ij = log(p_ij / p_ji),
         lambda_w = 0.5*logit_p_ij,
         lambda_l = -0.5*logit_p_ij)

df_2021 <- read_csv("data/Men 2021.csv", col_types = cols(Date = col_date(format = "%d/%m/%Y")))%>%
  filter(Tournament == "Wimbledon" & Round == "1st Round")%>%
  select(Winner, Loser, AvgW, AvgL)%>%
  mutate(p_ij = 1 / AvgW,
         p_ji = 1/ AvgL,
         total_prob = p_ij + p_ji) %>%
  mutate(p_ij = p_ij / total_prob,
         p_ji = p_ji / total_prob) %>%
  select(-total_prob) %>%
  mutate(logit_p_ij = log(p_ij / p_ji),
         lambda_w = 0.5*logit_p_ij,
         lambda_l = -0.5*logit_p_ij)
```


```{r}
atp18 <- read_csv("data/elo_ratings/atp18.csv")%>%
  rowwise()%>%
  mutate(myName = transform_name(Player))

write_csv(atp18, 'data/elo_ratings/atp18_myname.csv')

atp19 <- read_csv("data/elo_ratings/atp19.csv")%>%
  rowwise()%>%
  mutate(myName = transform_name(Player))

write_csv(atp19, 'data/elo_ratings/atp19_myname.csv')
```

The csv file for 2021 is messed up and needs some manual edits first, hence it is in Excel.

```{r}
atp21 <- atp21_mod <- read_excel("data/elo_ratings/atp21_mod.xlsx", 
    col_types = c("numeric", "text", "numeric", 
        "text", "text", "text", "text", "text", 
        "text"))%>%
  rowwise()%>%
  mutate(myName = transform_name(Player))

write_csv(atp21, 'data/elo_ratings/atp21_myname.csv')
```


```{r}
elo2018 <- df_2018%>%
  left_join(
    atp18%>%select(myName, Player, Elo), by = c("Winner" = "myName")
  )%>%
  rename("EloW" = "Elo", "PlayerW" = "Player")%>%
  left_join(
    atp18%>%select(myName, Player, Elo), by = c("Loser" = "myName")
  )%>%
  rename("EloL" = "Elo", "PlayerL" = "Player")

elo2019 <- df_2019%>%
  left_join(
    atp19%>%select(myName, Player, Elo), by = c("Winner" = "myName")
  )%>%
  rename("EloW" = "Elo", "PlayerW" = "Player")%>%
  left_join(
    atp19%>%select(myName, Player, Elo), by = c("Loser" = "myName")
  )%>%
  rename("EloL" = "Elo", "PlayerL" = "Player")

elo2021 <- df_2021%>%
  left_join(
    atp21%>%select(myName, Player, Elo), by = c("Winner" = "myName")
  )%>%
  rename("EloW" = "Elo", "PlayerW" = "Player")%>%
  left_join(
    atp21%>%select(myName, Player, Elo), by = c("Loser" = "myName")
  )%>%
  rename("EloL" = "Elo", "PlayerL" = "Player")
```


Get the missing names


```{r}
missings2018 <- c(
  elo2018%>%
  filter(is.na(PlayerW))%>%
  pull(Winner),
   elo2018%>%
  filter(is.na(PlayerL))%>%
  pull(Loser)
)%>%
  unique()

missings2019 <- c(
  elo2019%>%
  filter(is.na(PlayerW))%>%
  pull(Winner),
   elo2019%>%
  filter(is.na(PlayerL))%>%
  pull(Loser)
)%>%
  unique()

missings2021 <- c(
  elo2021%>%
  filter(is.na(PlayerW))%>%
  pull(Winner),
   elo2021%>%
  filter(is.na(PlayerL))%>%
  pull(Loser)
)%>%
  unique()
```



# Women

```{r}
df_2018 <- read_csv("data/Women 2018.csv", col_types = cols(Date = col_date(format = "%d/%m/%Y")))%>%
  filter(Tournament == "Wimbledon" & Round == "1st Round")%>%
  select(Winner, Loser, AvgW, AvgL)%>%
  mutate(p_ij = 1 / AvgW,
         p_ji = 1/ AvgL,
         total_prob = p_ij + p_ji) %>%
  mutate(p_ij = p_ij / total_prob,
         p_ji = p_ji / total_prob) %>%
  select(-total_prob) %>%
  mutate(logit_p_ij = log(p_ij / p_ji),
         lambda_w = 0.5*logit_p_ij,
         lambda_l = -0.5*logit_p_ij)

df_2019 <- read_csv("data/Women 2019.csv", col_types = cols(Date = col_date(format = "%d/%m/%Y")))%>%
  filter(Tournament == "Wimbledon" & Round == "1st Round")%>%
  select(Winner, Loser, AvgW, AvgL)%>%
  mutate(p_ij = 1 / AvgW,
         p_ji = 1/ AvgL,
         total_prob = p_ij + p_ji) %>%
  mutate(p_ij = p_ij / total_prob,
         p_ji = p_ji / total_prob) %>%
  select(-total_prob) %>%
  mutate(logit_p_ij = log(p_ij / p_ji),
         lambda_w = 0.5*logit_p_ij,
         lambda_l = -0.5*logit_p_ij)

df_2021 <- read_csv("data/Women 2021.csv", col_types = cols(Date = col_date(format = "%d/%m/%Y")))%>%
  filter(Tournament == "Wimbledon" & Round == "1st Round")%>%
  select(Winner, Loser, AvgW, AvgL)%>%
  mutate(p_ij = 1 / AvgW,
         p_ji = 1/ AvgL,
         total_prob = p_ij + p_ji) %>%
  mutate(p_ij = p_ij / total_prob,
         p_ji = p_ji / total_prob) %>%
  select(-total_prob) %>%
  mutate(logit_p_ij = log(p_ij / p_ji),
         lambda_w = 0.5*logit_p_ij,
         lambda_l = -0.5*logit_p_ij)
```


```{r}
wta18 <- read_csv("data/elo_ratings/wta18.csv")%>%
  rowwise()%>%
  mutate(myName = transform_name(Player))

wta19 <- read_csv("data/elo_ratings/wta19.csv")%>%
  rowwise()%>%
  mutate(myName = transform_name(Player))

write_csv(wta18, 'data/elo_ratings/wta18_myname.csv')
write_csv(wta19, 'data/elo_ratings/wta19_myname.csv')

```

The csv file for 2021 is messed up and needs some manual edits first, hence it is in Excel.

```{r, warning=FALSE}
wta21 <- read_excel("data/elo_ratings/wta21_mod.xlsx", 
    col_types = c("numeric", "text", "text", 
        "numeric", "text", "text", "text", 
        "text", "text", "text"))%>%
  rowwise()%>%
  mutate(myName = transform_name(Player))

write_csv(wta21, 'data/elo_ratings/wta21_myname.csv')
```


```{r}
elo2018 <- df_2018%>%
  left_join(
    wta18%>%select(myName, Player, Elo), by = c("Winner" = "myName")
  )%>%
  rename("EloW" = "Elo", "PlayerW" = "Player")%>%
  left_join(
    wta18%>%select(myName, Player, Elo), by = c("Loser" = "myName")
  )%>%
  rename("EloL" = "Elo", "PlayerL" = "Player")

elo2019 <- df_2019%>%
  left_join(
    wta19%>%select(myName, Player, Elo), by = c("Winner" = "myName")
  )%>%
  rename("EloW" = "Elo", "PlayerW" = "Player")%>%
  left_join(
    wta19%>%select(myName, Player, Elo), by = c("Loser" = "myName")
  )%>%
  rename("EloL" = "Elo", "PlayerL" = "Player")

elo2021 <- df_2021%>%
  left_join(
    wta21%>%select(myName, Player, Elo), by = c("Winner" = "myName")
  )%>%
  rename("EloW" = "Elo", "PlayerW" = "Player")%>%
  left_join(
    wta21%>%select(myName, Player, Elo), by = c("Loser" = "myName")
  )%>%
  rename("EloL" = "Elo", "PlayerL" = "Player")
```


Get the missing names


```{r}
missings2018 <- c(
  elo2018%>%
  filter(is.na(PlayerW))%>%
  pull(Winner),
   elo2018%>%
  filter(is.na(PlayerL))%>%
  pull(Loser)
)%>%
  unique()

missings2019 <- c(
  elo2019%>%
  filter(is.na(PlayerW))%>%
  pull(Winner),
   elo2019%>%
  filter(is.na(PlayerL))%>%
  pull(Loser)
)%>%
  unique()

missings2021 <- c(
  elo2021%>%
  filter(is.na(PlayerW))%>%
  pull(Winner),
   elo2021%>%
  filter(is.na(PlayerL))%>%
  pull(Loser)
)%>%
  unique()
```




# 2022


```{r}
transform_name <- function(name){
  splitName <- str_split(name, " ")[[1]]
  
  lastName <- tail(splitName, n=1)%>%str_to_title()
  firstName <- head(splitName, -1)%>%substr(start = 1, stop = 1)%>%toupper()%>%paste(".",sep = "")%>%paste(collapse = "")
  
  return(paste(lastName,firstName))
}

# transform odds data and get initial, unscaled lambda
transform_odds <- function(filename) {
  read_csv(filename) %>%
    mutate(p_ij = 1 / odds1,
           p_ji = 1 / odds2,
           total_prob = p_ij + p_ji) %>%
    mutate(p_ij = p_ij / total_prob,
           p_ji = p_ji / total_prob,
           logit_p_ij = log(p_ij / p_ji),
           lambda_1 = 0.5 * logit_p_ij,
           lambda_2 = -lambda_1)%>%
    select(-total_prob)
}

```


```{r}
# Odds
odds_men <- transform_odds("data/odds/men1.csv")
odds_women <- transform_odds("data/odds/women1.csv")
# Elo
atp22 <- read_csv("data/elo_ratings/atp22.csv")%>%
  rowwise()%>%
  mutate(odds_name = transform_name(Player))
wta22 <- read_csv("data/elo_ratings/wta22.csv")%>%
  rowwise()%>%
  mutate(odds_name = transform_name(Player))
```

```{r}
mergedM <- odds_men%>%
  left_join(
    atp22%>%select(odds_name, Player, Elo), by = c("player1" = "odds_name")
  )%>%
  rename("elo1" = "Elo", "elo_name_1" = "Player")%>%
  left_join(
    atp22%>%select(odds_name, Player, Elo), by = c("player2" = "odds_name")
  )%>%
  rename("elo2" = "Elo", "elo_name_2" = "Player")%>%
  select(player1, player2, elo_name_1, elo_name_2, everything())


mergedW <- odds_women%>%
  left_join(
    wta22%>%select(odds_name, Player, Elo), by = c("player1" = "odds_name")
  )%>%
  rename("elo1" = "Elo", "elo_name_1" = "Player")%>%
  left_join(
    wta22%>%select(odds_name, Player, Elo), by = c("player2" = "odds_name")
  )%>%
  rename("elo2" = "Elo", "elo_name_2" = "Player")%>%
  select(player1, player2, elo_name_1, elo_name_2, everything())
```

```{r}
missingsM <- c(
  mergedM%>%
  filter(is.na(elo_name_1))%>%
  pull(player1),
   mergedM%>%
  filter(is.na(elo_name_2))%>%
  pull(player2)
)%>%
  unique()

missingsW <- c(
  mergedW%>%
  filter(is.na(elo_name_1))%>%
  pull(player1),
   mergedW%>%
  filter(is.na(elo_name_2))%>%
  pull(player2)
)%>%
  unique()
```

```{r}
atp22_adjusted <- atp22%>%
  mutate(odds_name = case_when(
    Player == "Alejandro Davidovich Fokina" ~ "Davidovich Fokina A.",
    Player == "Tomas Martin Etcheverry" ~ "Etcheverry T. M.",
    Player == "Jan Lennard Struff" ~ "Struff J-L.",
    Player == "Tim Van Rijthoven" ~ "Van Rijthoven T.",
    Player == "Felix Auger Aliassime" ~ "Auger-Aliassime F.",
    Player == "Roberto Carballes Baena" ~ "Carballes Baena R.",
    Player == "Alex De Minaur" ~ "De Minaur A.",
    Player == "Daniel Elahi Galan" ~ "Galan Riveros D. E.",
    Player == "Marc Andrea Huesler" ~ "Huesler M.",
    Player == "Alexei Popyrin" ~ "Popyrin Al.",
    Player == "Bernabe Zapata Miralles" ~ "Zapata Miralles B.",
    Player == "Pablo Andujar" ~ "Andujar-Alba P.",
    Player == "Soon Woo Kwon" ~ "Kwon S.",
    Player == "Pablo Carreno Busta" ~ "Carreno-Busta P.",
    Player == "Roberto Bautista Agut" ~ "Bautista-Agut R.",
    Player == "Mackenzie Mcdonald" ~ "McDonald M.",
    Player == "Chun Hsin Tseng" ~ "Tseng C. H.",
    Player == "Botic Van De Zandschulp" ~ "Van De Zandschulp B.",
    !is.na(odds_name) ~ odds_name
  ))


wta22_adjusted <- wta22%>%
  mutate(odds_name = case_when(
    Player == "Jodie Anna Burrage" ~ "Burrage J.",
    Player == "Ylena In Albon" ~ "In-Albon Y.",
    Player == "Danielle Collins" ~ "Collins D. R.",
    Player == "Alison Van Uytvanck" ~ "Van Uytvanck A.",
    Player == "Irina Maria Bara" ~ "Bara Irina M.",
    Player == "Sara Sorribes Tormo" ~ "Sorribes Tormo S.",
    Player == "Jil Teichmann" ~ "Teichmann J. B.",
    Player == "Xin Yu Wang" ~ "Wang Xin.",
    Player == "Yuriko Miyazaki" ~ "Miyazaki L.",
    Player == "Beatriz Haddad Maia" ~ "Haddad Maia B.",
    Player == "Xiyu Wang" ~ "Wang Xiy.",
    Player == "Fernanda Contreras Gomez" ~ "Contreras Gomez F.",
    Player == "Irina Camelia Begu" ~ "Begu I.",
    Player == "Nastasja Mariana Schunk" ~ "Schunk N. M.",
    Player == "Elena Gabriela Ruse" ~ "Ruse G.",
    Player == "Nuria Parrizas Diaz" ~ "Parrizas-Diaz N.",
    Player == "Karolina Pliskova" ~ "Pliskova Ka.",
    Player == "Anna Karolina Schmiedlova" ~ "Schmiedlova A. K.",
    Player == "Christina Mchale" ~ "McHale C.",
    Player == "Daria Gavrilova" ~ "Saville D.",
    !is.na(odds_name) ~ odds_name
  ))


no_elo <- c(
  "Wickmayer Y.", "Kartal S.", "Williams S.", "Hives Z.",
"Yanina Wickmayer (not ok)", "Sonay Kartal (not ok)", "Serena Williams (ok)", "Zoe Hives (not ok)")
```

Sanity check

```{r}
mergedM_adjusted <- odds_men%>%
  left_join(
    atp22_adjusted%>%select(odds_name, Player, Elo), by = c("player1" = "odds_name")
  )%>%
  rename("elo1" = "Elo", "elo_name_1" = "Player")%>%
  left_join(
    atp22_adjusted%>%select(odds_name, Player, Elo), by = c("player2" = "odds_name")
  )%>%
  rename("elo2" = "Elo", "elo_name_2" = "Player")%>%
  select(player1, player2, elo_name_1, elo_name_2, everything())

```


```{r}
mergedW_adjusted <- odds_women%>%
  left_join(
    wta22_adjusted%>%select(odds_name, Player, Elo), by = c("player1" = "odds_name")
  )%>%
  rename("elo1" = "Elo", "elo_name_1" = "Player")%>%
  left_join(
    wta22_adjusted%>%select(odds_name, Player, Elo), by = c("player2" = "odds_name")
  )%>%
  rename("elo2" = "Elo", "elo_name_2" = "Player")%>%
  select(player1, player2, elo_name_1, elo_name_2, everything())
```


Save files

```{r}
write.csv(atp22_adjusted, file('data/elo_ratings/atp22_name_adjusted.csv', encoding = 'UTF-8'),
          row.names = FALSE)
write.csv(wta22_adjusted, file('data/elo_ratings/wta22_name_adjusted.csv', encoding = 'UTF-8'),
          row.names = FALSE)
```



